---
title: "Road accident cause analysis and accident hotspot prediction with Machine Learning"
subtitle: WQD7004 Programming for Data Science Group Project - Group 14
author: "Mithirendra Maniam"
date: "2023-11-02"
output: html_document
---
---

## 1. PROJECT INTRODUCTION

```{r, echo=FALSE, out.width="50%"}
# Define variable containing url
url <- "https://res.cloudinary.com/dmeglmbmj/image/upload/v1699100625/Traffic_Crash_d1p0rh.jpg"
```

<center><img src="`r url`" width="500" /></center>  

Road accidents are a major concern all over the world, since they cause casualties, injuries, and fatalities each year. Accidents also cause significant economic losses. The 2018 report of the World Health Organization states that more than 1.35 million people die each year from causes related to road accidents. An additional 20-50 million are injured or disabled [1]. If factors that cause accidents can be better understood and predicted, it might be possible to take measures to mitigate the damages and its severity [2].

---

## 2. PROJECT RESEARCH QUESTIONS
Many factors are responsible for causing road accidents. How to reduce the occurrence of fatal traffic accidents and improve road safety has been a significant problem for both governments and research institutions. Knowing what the influential factors are and how they affect the accidents can help better understand the cause-effect behind. This is beneficial to improve the estimation of the accident severity and preparation of countermeasures.[3]

The purpose of this project is to predict road accident hotspots and analyse non-human factors causing accidents, using traffic crash data from 2011 to 2023 from the New Zealand governments open data portal (data.gov.nz). Here are research questions that the project is aiming to answer:

1. What are the major accident areas and vehicle types in New Zealand?
2. Are there any non-human factors that cause more severe accidents? 
3. Can the data predict future severity of road accidents based on past data?  

---

## 3. PROJECT RESEARCH OBJECTIVES
The research objectives for the project are:

1. Identify major accident areas and vehicle types involved.
2. Identify non-human variables that may cause and increase the likelihood of severe accidents, including high fatalities.
3. Develop machine learning models to predict future road accident severity.  

---

## 4. LITERATURE REVIEW
Given the significant impact to society and mortality rates, accident prediction has been extensively studied with many different models. From a methodology perspective, there are many that have been explored. In 2019, Herbert et. al. [4] used a balanced random forest algorithm to study the accidents that occurred in Montreal. Overall, the algorithms predicted 85 percent of Montreal incidents, with a false positive rate (FPR) of 13%. Another study in 2019 on a GIS-based data mining methodology framework to analyze the influential factors of fatal traffic accidents showed that XGBoost obtained the highest modeling accuracy. [3] Fiorentini in 2020 gave outcomes from a random undersampling majority-class (RUMC) based models provide an enhancement in the reliability of the classifiers for detecting fatal crashes and those causing injury. Indeed, in imbalanced models, which showed that for the RUMC-based models, it spans from 52.5% (RUMC-based logistic regression) to 57.2% (RUMC-based k-nearest neighbor). Organizations and decision-makers could make use of RUMC and machine learning algorithms in predicting the severity of a crash occurrence, managing the present, and planning the future of their works [6].

There are many factors that cause traffic accidents. Previous research showed the most often factor contributing to the accident occurrence are human factors, with driver inattention being the highest, which could be caused by several causes as e.g.distraction, overloading attention, monotonous driving, etc.[5]. However, non-human factors have also caused accidents. One study by Jalilian et. al in 2019 [7] showed a significant relationship between fatal RTAs and factors such as; the sort of the road, the hindered visibility, the location of the accident, the accidents’ place, the climate, and lighting of the day (P<0.05). When it was cloudy, the chance was 2.60 times more than when was clear (P<0.05). But the sample size used was small, with only 2314 accidents dataset examined.

---

## 5. DATA COLLECTION

The dataset comes from the Waka Kotahi Crash Analysis System (CAS), which records all traffic crashes reported to data.gov.nz by the NZ Police. CAS covers crashes on all New Zealand roadways or places where the public have legal access with a motor vehicle.  

The Dataset URL is as follows: 

1. [Dataset Website](https://catalogue.data.govt.nz/dataset/crash-analysis-system-cas-data5)  
2. [Actual Dataset URL](https://opendata-nzta.opendata.arcgis.com/datasets/NZTA::crash-analysis-system-cas-data-1.csv?where=1=1&outSR=%7B%22latestWkid%22%3A2193%2C%22wkid%22%3A2193%7D)  

The dataset was downloaded from data.gov.nz on 30 Oct 2023. As of 30 Oct 2023, data was available from economic year 1999/2000 to 2022/2023, and has 821,744 observations.


```{r}
# Read csv file that was downloaded
traffic_df <- read.csv("Crash_Analysis_System_(CAS)_data.csv")

```

Display the structure of dataset.

```{r}
# Structure of dataset
str(traffic_df)

```

---

## 6. DATA CLEANING (Sundeep / Yao Hong)
# Load the required libraries
library(dplyr)
# retrieve data
data <- read.csv("C:/Users/Administrator/Documents/GitHub/TrafficCrashesProject/Crash_Analysis_System_(CAS)_data.csv")

# Handle missing values: fill in missing values instead of deleting them
# For numeric data, use median padding
# For categorical data, use the most common category padding

for(col in names(data)){
  if(is.numeric(data[[col]])){
    data[[col]][is.na(data[[col]])] <- median(data[[col]], na.rm = TRUE)
  } else if(is.factor(data[[col]]) || is.character(data[[col]])){
    levels <- unique(data[[col]][!is.na(data[[col]])])
    data[[col]][is.na(data[[col]])] <- levels[which.max(tabulate(match(data[[col]], levels)))]
  }
}
# Handling outliers: using the box plot rule for numeric columns
data <- data %>%
  mutate(across(where(is.numeric), ~replace(.x, .x < (quantile(.x, .25, na.rm = TRUE) - 1.5 * IQR(.x, na.rm = TRUE)) | .x > (quantile(.x, .75, na.rm = TRUE) + 1.5 * IQR(.x, na.rm = TRUE)), NA)))


# Can choose how to handle these outliers, e.g. replace median with
data <- data %>%
  mutate(across(where(is.numeric), ~replace(.x, is.na(.x), median(.x, na.rm = TRUE))))

# Remove duplicate records
data <- distinct(data)

# Export cleaned data
write.csv(data, "C:/Users/Administrator/Documents/GitHub/TrafficCrashesProject/1/CAS_data_cleaned.csv", row.names = FALSE)


---

## 7. EXPLORATORY DATA ANALYSIS  (Sundeep / Yao Hong)

library(ggplot2)
library(dplyr)

data <- read.csv("C:/Users/Administrator/Documents/GitHub/TrafficCrashesProject/1/CAS_data_cleaned.csv")

# 1. Overview of data
str(data)

# 2. Summary of data
summary(data)

## 3. missing value checks
cat("缺失值总数:", sum(is.na(data)), "\n")

# 4. Univariate Analysis - Plotting Histograms
# Take 'areaUnitID' as an example, you can replace it with any numeric column
hist(data$areaUnitID, main="直方图 - areaUnitID", xlab="areaUnitID")

# Univariate analysis - plotting box plots
# Take 'areaUnitID' as an example

boxplot(data$areaUnitID, main="盒形图 - areaUnitID", ylab="areaUnitID")

# 5. Multivariate analysis
# Plotting a scatterplot of X and Y
plot(data$X, data$Y, main="散点图 - X vs Y", xlab="X", ylab="Y")

# Correlation matrix (only numerical columns are considered)
numeric_data <- data[, sapply(data, is.numeric)]
cor_matrix <- cor(numeric_data, use="complete.obs")
print(cor_matrix)


# 6. Subgroup analysis
# calculating the mean of 'areaUnitID' based on the 'urban' classification
group_means <- tapply(data$areaUnitID, data$urban, mean)
print(group_means)

# 7. Visualisation (using ggplot2)
# create a bar chart based on 'crashSeverity' and 'areaUnitID'
ggplot(data, aes(x=crashSeverity, y=areaUnitID)) + geom_bar(stat="identity") + 
  ggtitle("条形图 - crashSeverity vs areaUnitID") + xlab("crashSeverity") + ylab("areaUnitID")


---




## 7. DATA METHODOLOGY  (Wei Ven / Ain)
  

---

## 8. DATA INTERPRETATION  (Wei Ven / Ain)
  

---

## 9. RECOMMENDATIONS AND FUTURE  (Mithi)
  

---

## 10. CONCLUSION  (Mithi)
  

---

## 11. OPTIONAL - DATA PRODUCT (All)


---

## 12. REFERENCES  (All)

1. WHO|10 Facts on Global Road Safety. Accessed: Oct. 10, 2018. [Online].Available: http://www.who.int/features/factfiles/roadsafety/en/
2. Santos, D., Salas, J., Quaresma, P. Machine Learning Approaches to Traffic Analysis and Hotspot Prediction. *Computers* 2021, 10(12), 157; [https://doi.org/10.3390/computers10120157](https://doi.org/10.3390/computers10120157)
3. Mai, J., Ding, Y., Cheng, J. C. P., Tan, Y., Gan, V. J. L.,Zhang, J.C. Analyzing the Leading Causes of Traffic Fatalities
Using XGBoost and Grid-Based Analysis: A City Management Perspective. *IEEE Access* 2019, Vol 7. [10.1109/ACCESS.2019.2946401](10.1109/ACCESS.2019.2946401)
4. Hébert, A.; Guédon, T.; Glatard, T.; Jaumard, B. High-Resolution Road Vehicle Collision Prediction for the City of Montreal. *Proceedings of the 2019 IEEE International Conference on Big Data (Big Data), Los Angeles*, CA, USA, 9–12 December 2019.
5. Bucsuházya, K., Matuchováa, E., Zůvalaa, R., Moravcováa, P., Kostíkováa, M., Mikuleca, R. Human factors contributing to the road traffic accident occurrence. AIIT 2nd International Congress on Transport Infrastructure and Systems in a changing world
(TIS ROMA 2019), 23rd-24th September 2019, Rome, Italy.
6. Fiorentini, N., Losa, M. Handling Imbalanced Data in Road Crash Severity Prediction by Machine Learning Algorithms. *Infrastructures* 2020, 5(7), 61; [https://doi.org/10.3390/infrastructures507006](https://doi.org/10.3390/infrastructures507006)
7. Jalilian, M.M, Safarpour, H., Bazyar, J., Keykaleh, M.S., Malekyan, L., Khorshidi, A. Environmental Related Risk Factors to Road Traffic Accidents in Ilam, Iran. *Med Arch.* 2019 Jun; 73(3): 169–172.



